---
title: "Yellow-legged frog abundances"
author: "Mae Rennick"
output: 
  distill::distill_article:
    self_contained: false
    toc: true
    toc_float: true
    theme: yeti
    highlight: tango
    code_folding: hide
---

## Overview 

The goal of these taks are to create two data visualizations put together in a compunded figure using `cowplot()` by exploring anphibian abundance data recorded by the  the Sierra Lakes Inventory Project.

## Data Summary and citation
From the Environmental Data Initiative repository: “The Sierra Lakes Inventory Project (SLIP) was a research endeavor that ran from 1995-2002 and has supported research and management of Sierra Nevada aquatic ecosystems and their terrestrial interfaces. We described the physical characteristics of and surveyed aquatic communities for >8,000 lentic water bodies in the southern Sierra Nevada, including lakes, ponds, marshes, and meadows.


Knapp, R.A., C. Pavelka, E.E. Hegeman, and T.C. Smith. 2020. The Sierra Lakes Inventory Project: Non-Native fish and community composition of lakes and ponds in the Sierra Nevada, California ver 2. Environmental Data Initiative. https://doi.org/10.6073/pasta/d835832d7fd00d9e4466e44eea87fab3


## Setup

```{r setup, include=TRUE, warning= FALSE, message=FALSE}
knitr::opts_chunk$set(echo = TRUE, warning= FALSE, message=FALSE)

#load necessary packages

library(tidyverse)
library(here)
library(lubridate)
library(cowplot)


```

## Construction of Graph A

Data wrangling and plot: mountain yellow-legged frog (Rana muscosa) count each year across all water bodies, by life stage (excluding egg mass)

```{r}

## read in the data using here()

frog_data<- read_csv(here("_code", "frogs", "data", "sierra_amphibians.csv"))

RAMU_data<- frog_data %>% 
  filter(amphibian_species== "RAMU") %>% 
  filter(amphibian_life_stage != "EggMass") %>% 
  select(-lake_id) %>% 
  select(-amphibian_location) %>%
  mutate(survey_date= mdy(survey_date)) %>% 
  mutate(year = year(survey_date)) %>% 
  mutate(count= 1) %>% 
  group_by(year, amphibian_life_stage) %>% 
  mutate(total= sum(count)) %>% 
  ungroup() %>% 
  summarise(year, amphibian_life_stage, total)

p1<- ggplot(RAMU_data, aes(x= year, fill= amphibian_life_stage))+
  geom_histogram()+
  labs(y= "total yellow-legged frog counts",
       fill= "Amphibian Life Stage")+
  scale_fill_manual(values = c("dark green", "light green", "brown"))+
  theme_minimal()+ 
  theme(legend.position= "bottom", legend.title = element_blank())


```

## Construction of Graph B

Data wrangling and plot: containing total counts (over all years of the study) of combined adult and subadult endangered mountain yellow-legged frogs (Rana muscosa) observed in the 5 lakes with the greatest total observed counts


```{r}

RAMU_data_adult<- frog_data %>% 
  filter(amphibian_species== "RAMU") %>%
  filter(amphibian_life_stage== "SubAdult"|  amphibian_life_stage=="Adult") %>%
  mutate(count=1) %>% 
  group_by(lake_id) %>% 
  summarize(frog_count = n()) %>% 
  slice_max(frog_count, n = 5) %>% 
  arrange(-frog_count) %>% 
  mutate(lake= "Lake") %>% 
  relocate(lake) %>% 
  unite("lake_id", lake:lake_id, sep = " ")



p2<- ggplot(data = RAMU_data_adult, aes(x = fct_reorder(lake_id, frog_count), y = frog_count, fill= lake_id)) +
  geom_col() +
  labs(y = "total yellow-legged frog counts (adult + subadult)", x = "Lake ID") +
  theme_minimal()+
  scale_fill_manual(values=c("#011f4b", "#03396c", "#005b96", "#6497b1", "#b3cde0"))+
  theme(legend.position= 'none', axis.text.x = element_text(angle = 90))


```

## Combined Plots

```{r}

plot_grid(p1, p2, labels = c('A', 'B'), label_size = 12)

```

###### Figure 1: Abundances of Mountain yellow-legged frog (Rana muscosa) recorded by the The Sierra Lakes Inventory Project 1995-2002 (Knapp et al. 2020). A) counts of mountian yellow-legged frog each year across all water bodies, by life stage (excluding egg mass) B) Counts of adult and subadult mountian yellow-legged frog observed in the 5 lakes with the greatest total observed counts. All data from Knapp et al. 2020.
